\documentclass[11pt,nocut]{article}

\usepackage{../../latex_style/packages}
\usepackage{../../latex_style/notations}
\usepackage{fancyhdr}

\pagestyle{fancy}
\renewcommand{\headrulewidth}{1pt}
\fancyhead[R]{DSGA1024 - Fall 2021}
\fancyhead[L]{HW3 - Matrix Rank - due September 26, 2021 at 12pm}

\setcounter{section}{3}

\begin{document}

\input{../preamble_homeworks.tex}

\vspace{1cm}

\begin{problem}[2.5 points]
True or false? Give a reason if true and a counterexample if false.
\begin{enumerate}[label=\normalfont(\textbf{\alph*})]
  \item A square matrix with a column of zeros cannot be invertible.
  \item If every row of a matrix adds up to zero the matrix must be singular.
  \item If every column of a matrix adds up to zero the matrix must be singular.
  \item Every matrix with 1's down the main diagonal is invertible.
  \item If $A$ is invertible, then $A^{-1}$ and $A^2$ are invertible.
\end{enumerate}
\end{problem}

\begin{problem}[2.5 points]
    Let $ A = \begin{pmatrix}
        5a - 2 & 3a & 3a-3 \\
        -4a + 2 & -3a + 1 & -2a + 2 \\
        -4a + 2 & -3a & -2a + 3
    \end{pmatrix}
    $ for a real number $a$.
    \begin{enumerate}[label=\normalfont(\textbf{\alph*})]
        \item Determine the rank of $A$ for all values of $a$, remember that you can use linear operations on rows and columns.
        \item Determine a basis of the nullspace $Ker(A)$ and image space $Im(A)$ for $a = 0$.
    \end{enumerate}
\end{problem}

\begin{problem}[2 points]
    Let $ A = \begin{pmatrix}
        1 & 0\\
        2 & 3
    \end{pmatrix}
    $,  
    $$
    B = \begin{pmatrix}
        1 & 0 & 0 \\
        2 & 3 & 0 \\
        4 & 5 & 0
    \end{pmatrix}
    \text{ and }
    C = \begin{pmatrix}
        1 & 0 & 0 & 0 \\
        2 & 3 & 0 & 0 \\
        0 & 4 & 5 & 0 \\
        6 & 0 & 7 & 8
    \end{pmatrix}
    \text{ and }
    D = \begin{pmatrix}
        1 & 0 & 0 & 0 \\
        2 & 3 & 0 & 0 \\
        0 & 4 & 0 & 0 \\
        0 & 5 & 6 & 0
    \end{pmatrix}
    $$ 
    \begin{enumerate}[label=\normalfont(\textbf{\alph*})]
        \item Determine the ranks of $A$, $B$, $C$ and $D$ by reasoning on the dimension of the span of the columns of the matrices.
        \item Consider a generic lower-triangular matrix $M \in \R^{n\times n}$(meaning that the entries $M_{ij} = 0$ if $j>i$). Show that if all diagonal entries of $M$ are non-zero, $M$ is invertible. 
        \item Show that if one of the diagonal entries $M_{ii} = 0$, then the matrix is not invertible.
        \item Consider a generic upper-triangular matrix $M \in \R^{n\times n}$($M_{ij} = 0$ if $j<i$). At what condition is $M$ invertible?
    \end{enumerate}
\end{problem}

\begin{problem}[2 points]
    We will build a formal proof that for all $L\in R^{n \times m}$, $\rank(L) = \rank(L^\top)$.
    \begin{enumerate}[label=\normalfont(\textbf{\alph*})]
        \item Show that for any matrices $A\in \R^{n \times k}$, $B \in \R^{k \times m}$, $\rank(A) \geq \rank(AB)$.
        \item Given a matrix $L\in R^{n \times m}$, show that $\ker(L) = \ker(L^\top L)$ (hint: you may want to consider $y = Lx$).
        \item Using ({\bf a}) and ({\bf b}) Deduce that $\rank(L) = \rank(L^\top)$ (hint: use the rank-nullity theorem).
    \end{enumerate}

\end{problem}



\begin{problem}[$\star$] Let $A \in \R^{n\times n}$ given by 
    $$
    A = \begin{pmatrix}
        0 & 0 & \cdots & 0 & 1 \\
        0 & \cdots & \cdots & 1& 0 \\
        \vdots & & & \vdots\\
        0 & 1 & 0& \cdots & \vdots \\
        1 & 0 & \cdots & \cdots & 0
    \end{pmatrix}.
    $$
    Compute $A^k$ for all $k \in \mathbb{N}$.
\end{problem}

\end{document}
